---
title: "ggtree_basicExample"
author: "Roxana Zamudio"
output:
  html_document: default
  pdf_document: default
date: "09 March 2021"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Installing packages, for installing remove the #
```{r library}
# if (!requireNamespace("BiocManager", quietly = TRUE))
#     install.packages("BiocManager")
#  
# BiocManager::install("ggtree")

#install.packages("ape")
#install.packages("ggplot2")
#install.packages("phylotools")
#install.packages("dplyr")
#install.packages("tidytree")
#install.packages("phytools")

```
## Uploading packages
```{r library}
library(ape)
library(ggplot2)
library(ggtree)
library(phylotools)
library(dplyr)
library(tidytree)
library(phytools)
```

## Reading the tree and quick plot of the tree (this is a example for Acinetobacter baumannii [obtained from public database], in this case the tree was build using the free alignment approach )

```{r ggtree}
tree <- read.tree("/Users/zamudio/Downloads/ggtree_example/wholegenome_acitenobacter.nwk")
tree2<-midpoint.root(tree) #rotting the tree by the midpoint 
length(tree2$tip.label)
ggtree(tree2, ladderize=FALSE)
```
## Renaming the tip label using the file where there is a old (1st column) and new name (2nd column)

```{r ggtree}
data<-read.table("/Users/zamudio/Downloads/ggtree_example/rename_tip.label.txt", header = TRUE, sep = "\t")
head(data)
tree3 <- sub.taxa.label(tree2, data) # renaming the tip label
tree3$tip.label
```

## Importing the metadata

```{r ggtree}
meta_data = read.csv("/Users/zamudio/Downloads/ggtree_example/meta_data.csv", sep=",", row.names=1) 
head(meta_data)
meta_group = read.table("/Users/zamudio/Downloads/ggtree_example/group.txt", sep = "\t",header = TRUE, stringsAsFactors = FALSE)
head(meta_group)
```

## Plotting the tree and tip label with different colors

```{r ggtree}
p<-ggtree(tree3, ladderize=FALSE) + ylim(0,110) + xlim(0,0.01) + geom_treescale(x=0.01, y=1, offset=1)
p

#1st option
p2<-p %<+% meta_group + scale_fill_manual(values = c("CC2"="red","other"="grey")) +
  geom_tiplab(aes(fill = factor(group)), size=2,
              color = "black", # color for label font
              geom = "label",  # label, not text
              label.padding = unit(0.05, "lines"), # amount of padding around the labels
              label.size = 0,alpha=.5, align = TRUE)# size of label border
p2

#2nd option
p3<-p %<+% meta_group + geom_tiplab(aes(color = factor(group),fontface="bold"),align=TRUE, size=2.8) + scale_color_manual(values = c("CC2"="black","other"="purple"))                
p3
```


## Plotting the tree with the metadata

```{r ggtree}
hm <- gheatmap(p3,meta_data, offset = 0.0011, width=0.6, font.size=3.5, colnames_position= "top", colnames_angle = 90, colnames_offset_y = 0, hjust = 0) + scale_fill_manual(values=c("presence" = "black", "absence" = "gray"))      
hm

png(filename = "/Users/zamudio/Downloads/ggtree_example/WGtree_acitenobacter.png", height = 14, width = 16, units = "in", res = 300)
plot(hm)
dev.off()
```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
